### 2017-03-05 sources/Math/GeometricInterpretationOfCovarianceMatrix.Rmd
eigenvector interpretation

---
title: "Untitled"
author: "Daniel Frederico Lins Leite"
date: "5 March 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{R}
library(scales)
library(MASS)

generateData <- function (u, sdx, sdy){
  size <- 1000
  x0 <- rnorm(mean = u, sd = sdx, n = size)
  x1 <- rnorm(mean = u, sd = sdy, n = size)
  X <- matrix(x0, ncol = 1)
  cbind(X, matrix(x1, ncol = 1))
}

plotMatrix <- function(X){
  xdf <- data.frame(X)
  plot(xdf, ylim=c(-10, 10), asp=1, col = alpha("black",0.5))
}

plotEigenvectors <- function(X){
  C <- covarianceMatrix(X)
  eigenC <- eigen(C, symmetric = TRUE)
  dimensions <- eigenC$vectors %*% diag(sqrt(eigenC$values))
  arrows(0,0,dimensions[1,1],dimensions[2,1], lwd = 4, col = "red")
  arrows(0,0,dimensions[1,2],dimensions[2,2], lwd = 4, col = "blue")
}

rotationMatrix <- function(angle){
  matrix(c(cos(angle),sin(angle),-sin(angle),cos(angle)), nrow = 2, ncol= 2)
}

covarianceMatrix <- function(X){
  mu <- matrix(colMeans(X), nrow = 1)
  t(X)%*%X * (1/(nrow(X)-1)) - (t(mu)%*%mu)
}
```

# Theory

The eigenvectors of the covariance matrix should form a new basis perpendicular to variation of the data.

```{R}
C <- diag(1,2,2)
eigenC <- eigen(C)
dimensions <- eigenC$vectors %*% diag(eigenC$values)
C
eigenC
dimensions
```

# White Data

A white data (Gaussian with mean = 0 and Standard Deviation = 1) should have its eigenvector as near as possivle to the default basis.

```{R}
X <- generateData(0,1,1)
plotMatrix(X)
plotEigenvectors(X)
```

# Scaled Gaussian

A Scaled Gaussian should have one bigger eigenvector pointing to the variation of the data.

```{R}
X <- generateData(0, 4, 1)
plotMatrix(X)
plotEigenvectors(X)
```

# Scaled and Rotated Gaussian

The Scaled and Rotated Gaussian should have its eigenvectors rotated adn poiting to new variation direction.

```{R}
X <- generateData(0, 4, 1)
rot45 <- rotationMatrix(-pi/4)
X <- X %*% rot45
C <- covarianceMatrix(X)
eigenC <- eigen(C)
plotMatrix(X)
plotEigenvectors(X)
```

# Covariance Matrix is the Linear Transformation Squared

The Scaled and Rotated Gaussian can be generated from a White Data Gaussian using two Linear Transformations:
1 - Scale Transformation
2 - Rotation Transformation

```{R}
X <- generateData(0,1,1)
R <- rotationMatrix(-pi/4)
S <- matrix(c(4,0,0,1), nrow=2, ncol=2)
plotMatrix(X)
```

The complete transformation can be aggregated in the Matrix:

```{R}
T <- R%*%S
```

In thos case the covariance matrix is:

```{R}
#R is a orthonormal matrix so t(R) = ginv(R)
C <- T%*%t(T)
```

The final data is identical

```{R}
plotMatrix(X%*%S%*%R)
```

In this case its eigenvectors can be calculated as:

```{R}
eigenVectors <- t(R)%*%S
plotMatrix(X%*%S%*%R)
arrows(0,0,eigenVectors[1,1],eigenVectors[2,1], lwd = 4, col = "red")
arrows(0,0,eigenVectors[1,2],eigenVectors[2,2], lwd = 4, col = "blue")

# These matrices should be as near as zero as possible
eigenC$vectors - t(R)
diag(sqrt(eigenC$values), 2, 2) - S
```