### 2017-03-25 texts/philosophy/introolavo.tex
intro to olavos work

\documentclass[10pt,a4paper]{book}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\begin{document}
	\title{Introduction to Olavo de Carvalho´s Work}
	\author{Daniel Frederico Lins Leite}

	\maketitle

	\tableofcontents

	\chapter{Introduction}
	\section{Moral Basis for Philosophy}
	\subsection{Sincerity}
	\subsection{Love}
	\subsection{Science}
	\section{Idea versus Reality}
	\subsection{Reality}
	\subsection{Idea}
	\section{Philosopher Maturity}
	\subsection{Impression of Reality}
	\subsubsection{Senses}
	\subsubsection{Limited Structure of the Reality}
	\subsection{Unification}
	\subsubsection{Reality Unification}
	\subsubsection{Conscience Unification}
	\subsubsection{12 layers of personality}
	\subsubsection{Cognitive Parallax}
	\subsection{Expression of Reality}
	\subsubsection{4 modes of discourses}
	\subsubsection{Poetic}
	\subsubsection{Rhetoric}
	\subsubsection{Dialectic}
	\subsubsection{Logic}

	\chapter{Applications}
	\section{Philosophy}
	\subsection{Status Quo}
	\subsection{History of the History of Philosophy}
	\subsection{Greek Philosophy}
	\subsection{Christian Philosophy}
	\subsection{Muslim Philosophy}
	\subsection{Modern Philosophy}
	\subsection{Current Philosophy}
	\section{Politics}
	\subsection{Status Quo}
	\subsection{Inversion of the Evolution of the State}
	\subsection{Conscience of Transcendent}
	\subsection{Messianic Discourse}
	\subsection{Revolutionary Mentality}
	\section{Science}
	\subsection{Status Quo}
	\subsection{Inversion of Modern Cosmological Views}
\end{document}