### 2017-03-30 texts/dev/PapersILove.tex
development material

\documentclass{article}
\usepackage{hyperref}

\begin{document}

\title{Papers I Love}
\author{Daniel Frederico Lins Leite}

\maketitle

\tableofcontents

\section{Computer Science}

\subsection{Algorithms}

\subsubsection{Analysis}
\begin{enumerate}
	\item {Recursive Algorithms in Computer Science Courses: Fibonacci Numbers and Binomial Coefficients\\
\url{http://venus.cs.qc.edu/~waxman/cs211%20spring%202009/why%20is%20recursive%20fibonacci%20so%20slow.pdf}}
	\item {Binomial Coefﬁcient Computation: Recursion or Iteration?\\
\url{http://delab.csd.auth.gr/papers/SBI02m.pdf}}

\end{enumerate}

\subsubsection{Compression}
\begin{enumerate}
\item {Data Compression Using Long Common Strings\\
\url{http://www.cs.brandeis.edu/~dilant/cs175/%5BSiying-Dong%5D.pdf}}
\end{enumerate}

\subsubsection{Hash}
\begin{enumerate}
	\item {SHA-1 and the Strict Avalanche Criterion\\
\url{https://arxiv.org/pdf/1609.00616.pdf}}
\end{enumerate}

\subsubsection{Data Structures}
\begin{enumerate}
\item {Bitlist New Full-Text Index for Low Space Cost and Efficient Keyword Search\\
\url{http://www.vldb.org/pvldb/vol6/p1522-rao.pdf}}
\end{enumerate}

\subsubsection{Elections + Consensus}
\begin{enumerate}
	\item {Elections in a Distributed Computing System\\
\href{http://academic.research.microsoft.com/Publication/716253/elections-in-a-distributed-computing-system}{http://academic.research.microsoft.com/Publication/716253/elections-in-a-distributed-computing-system}\\
\url{http://homepage.cs.uiowa.edu/~ghosh/Bully.pdf}}
	\item {The Part-Time Parliament\\
\href{http://research.microsoft.com/en-us/um/people/lamport/pubs/lamport-paxos.pdf}{http://research.microsoft.com/en-us/um/people/lamport/pubs/lamport-paxos.pdf}}
	\item {In Search of an Understandable Consensus Algorithm\\
\href{https://ramcloud.atlassian.net/wiki/download/attachments/6586375/raft.pdf}{https://ramcloud.atlassian.net/wiki/download/attachments/6586375/raft.pdf}}
\end{enumerate}

\subsection{Architectures}

\subsubsection{Computer Architecture}
\begin{enumerate}
	\item {Quantifying the Cost of Context Switch\\
\url{http://www.cs.rochester.edu/u/cli/research/switch.pdf}}
	\item {What Every Programmer Should Know About Memory\\
\url{https://people.freebsd.org/~lstewart/articles/cpumemory.pdf}}
\end{enumerate}

\subsubsection{Multi Tenancy}
\begin{enumerate}
	\item {Enabling Multi-Tenancy an Industrial Experience Report\\
\href{http://swerl.tudelft.nl/twiki/pub/Main/TechnicalReports/TUD-SERG-2010-030.pdf}{http://swerl.tudelft.nl/twiki/pub/Main/TechnicalReports/TUD-SERG-2010-030.pdf}}
	\item {Multi-Tenant Saas Applications: Maintenance Dream or Nightmare\\
\href{http://swerl.tudelft.nl/twiki/pub/Main/TechnicalReports/TUD-SERG-2010-031.pdf}{http://swerl.tudelft.nl/twiki/pub/Main/TechnicalReports/TUD-SERG-2010-031.pdf}}
	\item {Towards an Elastic and Autonomic Multitenant Database\\ \url{http://research.microsoft.com/en-us/um/people/srikanth/netdb11/netdb11papers/netdb11-final8.pdf}}
\end{enumerate}

\subsubsection{REST}
\begin{enumerate}
	\item {Architectural Styles and the Design of Network-Based Software Architectures\\
\href{http://academic.research.microsoft.com/Publication/1309313/architectural-styles-and-the-design-of-network-
	based-software-architectures}{http://academic.research.microsoft.com/Publication/1309313/architectural-styles-and-the-design-of-network-
	based-software-architectures}\\
\url{http://www.ics.uci.edu/~fielding/pubs/dissertation/top.htm}}
\end{enumerate}

\subsubsection{SEDA}
\begin{enumerate}
	\item {An Architecture for Highly Concurrent, Well-Conditioned Internet Services\\
\href{http://academic.research.microsoft.com/Publication/112151/seda-an-architecture-for-well-conditioned-scalable-internet-services}{http://academic.research.microsoft.com/Publication/112151/seda-an-architecture-for-well-conditioned-scalable-internet-services}\\
\url{http://www.eecs.harvard.edu/~mdw/papers/mdw-phdthesis.pdf}}
\end{enumerate}

\subsubsection{Servers}
\begin{enumerate}
	\item {Flash an Efficient and Portable Web Server\\
\url{https://www.usenix.org/event/usenix99/full_papers/pai/pai.pdf}}
\end{enumerate}

\subsubsection{Other Architectures}
\begin{enumerate}
	\item {The Monad Manifesto\\
\url{http://www.jsnover.com/Docs/MonadManifesto.pdf}}
	\item {The Hla Tutorial\\
\url{http://www.pitch.se/hlatutorial}}
\end{enumerate}

\subsubsection{Patterns}
\begin{enumerate}
	\item {Active Object: An Object Behavioral Pattern for Concurrent Programming\\
\url{http://www.cs.wustl.edu/~schmidt/PDF/Act-Obj.pdf}}
	\item {Plop Half-Sync/half-Async: An Architectural Pattern for Efficient and Well-Structured Concurrent I/o\\
\url{http://www.cs.wustl.edu/~schmidt/PDF/PLoP-95.pdf}}
\end{enumerate}

\subsubsection{Overlay Networks}
\begin{enumerate}
	\item {Architectures for an Event Notification Service Scalable to Wide-Area Networks\\
\href{http://academic.research.microsoft.com/Publication/314658/architectures-for-an-event-notification-service-scalable-to-wide-area-networks}{http://academic.research.microsoft.com/Publication/314658/architectures-for-an-event-notification-service-scalable-to-wide-area-networks}\\
\url{http://www.inf.usi.ch/carzaniga/papers/phd_thesis.pdf}}
\end{enumerate}

\subsubsection{Distributed Systems}
\begin{enumerate}
	\item {Time, Clocks and the Ordering of Events in a Distributed System\\		
\href{http://academic.research.microsoft.com/Publication/775212/time-clocks-and-the-ordering-of-events-in-a-distributed-system}{http://academic.research.microsoft.com/Publication/775212/time-clocks-and-the-ordering-of-events-in-a-distributed-system}\\
\href{http://research.microsoft.com/en-us/um/people/lamport/pubs/pubs.html#time-clocks}{http://research.microsoft.com/en-us/um/people/lamport/pubs/pubs.html\#time-clocks}\\
\href{http://research.microsoft.com/en-us/um/people/lamport/pubs/time-clocks.pdf}{http://research.microsoft.com/en-us/um/people/lamport/pubs/time-clocks.pdf}}
	\item {Distributed Snapshots: Determining Global States of Distributed Systems\\
\href{http://academic.research.microsoft.com/Publication/803548/distributed-snapshots-determining-global-states-of-distributed-systems}{http://academic.research.microsoft.com/Publication/803548/distributed-snapshots-determining-global-states-of-distributed-systems}\\
\href{http://research.microsoft.com/en-us/um/people/lamport/pubs/pubs.html#chandy}{http://research.microsoft.com/en-us/um/people/lamport/pubs/pubs.html\#chandy}\\
\href{http://research.microsoft.com/en-us/um/people/lamport/pubs/chandy.pdf}{http://research.microsoft.com/en-us/um/people/lamport/pubs/chandy.pdf}}
	\item {Your Coffee Shop Doesn’t Use Two-Phase Commit\\
\href{http://www.enterpriseintegrationpatterns.com/docs/IEEE_Software_Design_2PC.pdf}{http://www.enterpriseintegrationpatterns.com/docs/IEEE\_Software\_Design\_2PC.pdf}}
	\item {A Brief History of Process Algebra\\
\url{http://alexandria.tue.nl/extra1/wskrap/publichtml/200402.pdf}}
	\item {Life Beyond Distributed Transactions: An Apostate’s Opinion\\
\url{http://www.ics.uci.edu/~cs223/papers/cidr07p15.pdf}}
\end{enumerate}

\subsubsection{Event Based Architecture}
\begin{enumerate}
	\item {Design of a Scalable Event Notification Service Interface and Architecture\\
\href{http://academic.research.microsoft.com/Publication/312680/design-of-a-scalable-event-notification-service-interface-and-architecture}{http://academic.research.microsoft.com/Publication/312680/design-of-a-scalable-event-notification-service-interface-and-architecture}\\
\url{http://www.inf.usi.ch/carzaniga/papers/CU-CS-863-98.pdf}}
	\item {Fast Forwarding for Content-Based Networking\\
\href{http://academic.research.microsoft.com/Publication/7217/fast-forwarding-for-content-based-networking}{http://academic.research.microsoft.com/Publication/7217/fast-forwarding-for-content-based-networking}\\
\href{http://www.inf.usi.ch/carzaniga/papers/cucs-922-01-r1.pdf}{http://www.inf.usi.ch/carzaniga/papers/cucs-922-01-r1.pdf}}
	\item {Real-Time Modelling of Dds for Event-Driven Applications\\
\url{http://www.ctr.unican.es/publications/hpt-jjg-2012a.pdf}}
\end{enumerate}

\subsubsection{Resiliency}
\begin{enumerate}
	\item {Adaptive Overload Control for Busy Internet Servers\\
\href{http://academic.research.microsoft.com/Publication/634136/adaptive-overload-control-for-busy-internet-servers}{http://academic.research.microsoft.com/Publication/634136/adaptive-overload-control-for-busy-internet-servers}\\
\url{http://www.eecs.harvard.edu/~mdw/papers/control-usits03.pdf}}
\end{enumerate}

\subsection{Programming Paradigms}
\subsubsection{Object Oriented}
\begin{enumerate}
	\item {A Theory of Objects\\
\url{http://academic.research.microsoft.com/Publication/1354440/a-theory-of-objects}\\
\url{http://lucacardelli.name/Talks/1997-06\%20A\%20Theory\%20of\%20Object\%20(ECOOP\%20Tutorial).pdf}}
	\item {Traits: Composable Units of Behaviour\\
\url{http://scg.unibe.ch/archive/papers/Scha03aTraits.pdf}}
	\item {Applying Traits to the Smalltalk Collection Hierarchy\\
\url{http://www.researchgate.net/publication/2564879_Applying_Traits_to_the_Smalltalk_Collection_Hierarchy}}
	\item {A Laboratory for Teaching Object-Oriented Thinking\\
\url{http://www.inf.ed.ac.uk/teaching/courses/seoc/2007_2008/resources/CRC_OOthinking.pdf}}
\end{enumerate}
\subsubsection{Double Dispatch}
\begin{enumerate}
	\item {Design and evaluation of C++ open multi-methods\\
\url{https://parasol.tamu.edu/~yuriys/papers/OMM10.pdf}}
\end{enumerate}

\subsection{Database}
\begin{enumerate}
	\item {The Ubiquitous B-Tree\\
\url{http://people.cs.aau.dk/~simas/aalg06/UbiquitBtree.pdf}}
	\item {Generalized Search Trees for Database Systems\\
\url{http://db.cs.berkeley.edu/papers/vldb95-gist.pdf}}
	\item {Concurrency and Recovery in Generalized Search TreeS\\
\url{http://db.cs.berkeley.edu/papers/sigmod97-gist.pdf}}
	\item {Data Cube: A Relational Aggregation Operator Generalizing Group-By, Cross-Tab, and Sub-Totals\\
\url{http://research.microsoft.com/pubs/69578/tr-95-22.pdf}}
	\item {Query Optimization in Microsoft Sql Server PDW\\
\href{http://academic.research.microsoft.com/Publication/56916436/query-optimization-in-microsoft-sql-server-pdw}{http://academic.research.microsoft.com/Publication/56916436/query-optimization-in-microsoft-sql-server-pdw}}
	\item {Druid: A Real-Time Analytical Data Store
\url{http://static.druid.io/docs/druid.pdf}}
	\item {Map-Reduce: Simplified Dataprocessing on Large Clusters\\
\url{http://static.googleusercontent.com/media/research.google.com/en/us/archive/mapreduce-
	osdi04.pdf}}
	\item {Google’s Mapreduce Programming Model — Revisited\\
\url{http://www.idt.mdh.se/kurser/cd5100/ht06/MapReduce/Ralf-Laemmel-paper/paper.pdf}}
	\item {Cassandra - a Decentralized Structured Storage System\\
\url{http://www.cs.cornell.edu/projects/ladis2009/papers/lakshman-ladis2009.pdf}}
	\item {Bigtable: A Distributed Storage System for Structured Data\\
\url{http://static.googleusercontent.com/media/research.google.com/en//archive/bigtable-osdi06.pdf}}
	\item {Dynamo: Amazon’s Highly Available Key-Value Store\\
\url{http://s3.amazonaws.com/AllThingsDistributed/sosp/amazon-dynamo-sosp2007.pdf}}
	\item {Solving Big Data Challenges for Enterprise Application Performance Management\\
\url{http://vldb.org/pvldb/vol5/p1724_tilmannrabl_vldb2012.pdf}}
\end{enumerate}

\subsection{Data Fusion}
\begin{enumerate}
	\item {A Generic Architecture for Fusion-Based Intrusion Detection Systems\\
\url{https://rcdeboer.home.xs4all.nl/rcdb_thesis.pdf}}
\end{enumerate}

\section{Mathematics}
\subsection{Real Analysis}
\begin{enumerate}
	\item {COISAS QUE O LUÍS PRECISA APRENDER\\
\url{http://www.todasasconfiguracoes.com/wp-content/uploads/2012/04/luis.pdf}}
\end{enumerate}
\subsection {Statistics}
\begin{enumerate}
	\item {A Note on the Generation of Random Normal Deviates\\
\url{http://projecteuclid.org/euclid.aoms/1177706645}}
	\item {Tidy Data\\
\url{http://vita.had.co.nz/papers/tidy-data.pdf}}
	\item {A Tutorial on Principal Component Analysis - Derivation, Discussion and Singular Value Decomposition\\
\url{https://www.cs.princeton.edu/picasso/mats/PCA-Tutorial-Intuition_jp.pdf}}
	\item{An introduction to ROC analysis\\
\url{https://ccrma.stanford.edu/workshops/mir2009/references/ROCintro.pdf}}
\end{enumerate}

\section{Economy}
\subsection {Political Economy}
\subsubsection{Taxes}
\begin{enumerate}
	\item {The Laffer Curve Past, Present, and Future\\
\url{http://s3.amazonaws.com/thf_media/2004/pdf/bg1765.pdf}}
	\item {Dynamic Revenue Estimation\\
\url{https://ideas.repec.org/a/aea/jecper/v10y1996i1p141-57.html}}
	\item {Dynamic Scoring an Introduction to the Issues\\
\url{https://www.aeaweb.org/annual_mtg_papers/2005/0107_1430_1304.pdf}}
\end{enumerate}

\end{document}