### 2017-04-29 texts/dev/DesignPatterns.tex

\documentclass[10pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\begin{document}

\section{Object Oriented Patterns}
\subsection{GOF Patterns}
\subsubsection{Bridge}
\begin{enumerate}
	\item {Design Patterns for Generic Programming in C++\\ \url{https://www.lrde.epita.fr/dload/papers/coots01.html}}
	\item {Generic programming in POOMA and PETE\\ \url{https://www.researchgate.net/publication/221024955_Generic_programming_in_POOMA_and_PETE}}
\end{enumerate}

\section{Functional Parsers}
\subsection{Monoid}
\begin{enumerate}
	\item {\url{https://fsharpforfunandprofit.com/posts/monoids-without-tears/}}
\end{enumerate}

\subsection{Monad}
\subsubsection{Parsers Combinators}
\begin{enumerate}
	\item {\url{https://fsharpforfunandprofit.com/posts/understanding-parser-combinators/}}
\end{enumerate}

\end{document}