### 2018-02-17 courses/johnhopkins-datascience/london.rmd

---
title: "London Tube"
author: "Daniel Frederico Lins Leite"
date: "21 April 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
library(jsonlite)
library(RCurl)
library(dplyr)
```

## Realtime London Underground Position

page created at 2017-04-21 by Daniel Leite

```{r}
raw_data <- getURL("http://traintimes.org.uk/map/tube/data/london.json")

data <- fromJSON(raw_data, flatten = TRUE)
trains <- data$trains
trains["lat"] <- sapply(data$trains$point, function(x) x[1])
trains["long"] <- sapply(data$trains$point, function(x) x[2])

subicon <- makeIcon("http://cdn2.iconfinder.com/data/icons/circle-icons-1/64/train-32.png", 32, 32)
trains %>% 
  leaflet() %>%
  addTiles() %>%
  addMarkers(popup = trains$title, icon = subicon)
```