### 2018-05-06 books/FundamentalsMatrixComputations/exercise.01.003.025.tex
Fundamentals of Matrix Computations

\documentclass{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{xcolor}
    \begin{document}
    Write down the numbers $1, 2, 3, ..., n-1$ on a line. Immediately below 
   these, write down the numbers $n - 1, n - 2, n - 3,..., 1$. Add these numbers up by 
   summing column-wise first. Conclude that 
   $$2*(1 + 2 + ... + n - 1) = n(n - 1)$$

   \begin{align*}
	A & = 1+2+3+...+(n-1)\\
	B &= (n-1) + (n-2) + (n-3) + ... + 1\\
	A + B &= [1 + (n - 1)] + [2 + (n-2)] + ... + [(n-1) + 1]\\
	A + B &= \sum_{i=1}^{n-1}{[i+(n-i)]}\\
	A + B &= \sum_{i=1}^{n-1}{n}\\
	A + B &= n*\sum_{i=1}^{n-1}{1}\\	
	A + B &= n*(n-1)\\
   \end{align*}

    \end{document}