### 2018-05-07 books/FundamentalsMatrixComputations/Exercise.01.04.054.tex
Fundamentals of Computation Matrix - Cholesky Proof - preliminary exercises

\documentclass{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\begin{document}
	Proposition 1.4.53\\
	Let A be "Positive Definite", and consider a partition 
	$$A = \begin{Bmatrix}
	A_{11}&A_{12}\\
	A_{21}&A_{22}\\
	\end{Bmatrix}$$
	in which $A_{11}$ and $A_{22}$ are square. Then $A_{11}$ and $A_{22}$ are Positive Definite.\\ 
	Exercise 1.4.54\\
	Prove Proposition 1.4.53. As in the previous exercise, do not use the Cholesky 
	decomposition in your proof; use the fact that $x^TAx > 0$ for all nonzero x.\\
	\\
	Proof:\\
	\\
	\begin{align*}
		x^TAx = \begin{Bmatrix} x1^T & 0\end{Bmatrix}*\begin{Bmatrix}
			A_{11} & A_{12}\\
			A_{21} & A_{22}
		\end{Bmatrix}*\begin{Bmatrix} x1 \\ 0\end{Bmatrix}\\
		x^TAx = \begin{Bmatrix}
			x_1^T*A_{11} & x_1^T*A_{22}
		\end{Bmatrix}*\begin{Bmatrix} x1 \\ 0\end{Bmatrix}\\		
		x^TAx = x_1^T*A_{11}*x_1
	\end{align*}
	But given that A is "Positive Definite" it is necessary that $x^TAx >0$, but this imply that it is also necessary that $x^TA_{11}x > 0$.
	$$\square$$	
\end{document}