### 2018-05-13 books/FundamentalsMatrixComputations/Exercise.01.005.009.tex
Fundamentals of Computation Matrix - Banded backward substitution complexity

\documentclass{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{pgfplots}
\begin{document}
	Exercise 1.5.9\\
	\\
	Let $R$ be an $n \times n$ upper-triangular matrix with semiband width $s$. Show that the system $Rx = y$ can be solved by back substitution in about $2ns$ flops. An analogous result holds for lower-triangular systems.\\
	\\
	\begin{align*}
	\begin{bmatrix}
	a_{11} & a_{12} & ...    & a_{1s} & 0      & ...    & 0\\
	0      & a_{22} & a_{23} & ...    & a_{2s} & ...    & 0\\
	0      & 0      & a_{33} & a_{34} & ...    & ...    & 0\\
	\vdots & \vdots & \vdots & \vdots & \ddots & \vdots & \vdots\\
	0      & 0      & 0      & 0      & 0      & 0      & a_{nn}
	\end{bmatrix} \begin{bmatrix}
	x_1\\x_2\\x_3\\ \vdots \\ x_n
	\end{bmatrix} = \begin{bmatrix}
	b_1\\b_2\\b_3\\ \vdots \\ b_n
	\end{bmatrix}
	\end{align*}
	\\
	Proof:\\
	\\
	First step:\\
	\\
	$$x_n =  b_n / a_{nn}$$
	\\
	1 - flop, 2 - reads, 1 - write\\
	\\
	Next step would be:\\
	\\	
	$$x_{i} = (b_i - \sum_{j=i+1}^{n}(a_{ij}*x_{j})) / a_{ii}$$
	\\
	but, we now that from $a_{i,j+s+1}$ to $a_{in}$, all entries are zero because the matrix $A$ is banded with a semi-band of $s$. This will reflect in the sumation, so, in this case:
	\begin{align*}
		\sum_{j=i+1}^{n}(a_{ij}*x_{j}) &= \sum_{j=i+1}^{i+s}(a_{ij}*x_{j}) + \sum_{j=i+s+1}^{n}(0*x_{j})\\
		&= \sum_{j=i+1}^{i+s}(a_{ij}*x_{j})\\
	\end{align*}
	This will generate $(i+s) - (i+1) + 1 = i+s-i-1+1 = s$ iterations per line. Each iteration of the summation will perform:\\
	1 flop, 3 reads (we will read the accumulator), 1 write (updating the accumulator). Totalling:
	$s$ flops, $3s$ reads, $s$ writes.\\
	Plugging back in the original expression for line $i$ we have:
	(+1)Read $b_i$, (+1)flop $b_i - \text{acc}$, (+1)read $a_{ii}$, (+1)flop $/a_{ii}$, (+1)write $x_i$.
	In total we have:
	$s+2$ flops, $3s+2$ reads, $s+1$ writes for each line (worst case).
	Now we can calculate the cost for the whole matrix, just assuming (incorrectly) that each line takes the worst case and arrive at:\\
	$n*(s+2)=ns+2n$ flops,\\
	$n*(3s+2)=3ns+2n$ reads,\\
	$n*(s+1)=ns+n$ writes.\\
	In this banded matrix scenario s << n and it is not an absurd to approximate $s ~ 2$ that would gives us the $2ns$ flops of the exercise.
	$$\square$$
	\\\\
	But to better understand the cost of the algorithm we will need to reference Agner4
	\begin{verbatim}
	Latency:
	This is the delay that the instruction generates in a dependency chain.
	The numbers are 	minimum values. Cache misses, misalignment, and 
	exceptions may increase the clock counts considerably. Where 
	hyperthreading is enabled, the use of the same execution units in the 
	other thread leads to inferior performance. Denormal numbers, NAN's 
	and infinity do not increase the latency. The time unit used is core 
	clock cycles, not the reference	clock cycles given by the time stamp 
	counter.

	Throughput:
	The average number of core clock cycles per instruction for a series of 
	independent instructions of the same kind in the same thread.

	Instruction Operand 	Latency Throughput  
	FADD        Register    3       1
	FSUB        Register    3       1
	FADD        Memory      5       1
	FSUB        Memory      5       1	
	FMUL        Register    5       1
	FDIV        Register    16      5
	FDIV        Memory      16      5
	\end{verbatim}	
	\url{http://www.agner.org/optimize/instruction_tables.pdf}
	\\
	\\
	What we must analyse is that these are best-case scenarios. As long we have cache misses these values can,and will increase enormously. We can see the cost of cache misses, for example, at:\\
	\begin{verbatim}
		Level               Fastest Latency
		First Level Data    4 cycles
		Second Level        12 cycles
		Third Level         44 cycles
	\end{verbatim}
	\url{https://www.intel.co.uk/content/dam/www/public/us/en/documents/manuals/64-ia-32-architectures-optimization-manual.pdf}\\
	\\
	Tehese, again, are the best scenario. And we are not event touching the main system meory yet.\\
	Memory access is very difficult to stimate, but we will use the value used at:\\
	\begin{verbatim}
		Data Source                                 Latency
		L3 CACHE hit, line unshared                 ~40 cycles
		L3 CACHE hit, shared line in another core   ~65 cycles
		L3 CACHE hit, modified in another core      ~75 cycles
		remote L3 CACHE                             ~100-300 cycles
		Local Dram                                  ~60 ns
		Remote Dram                                 ~100 ns 
	\end{verbatim}
	\url{https://software.intel.com/sites/products/collateral/hpc/vtune/performance_analysis_guide.pdf}\\
	\\\\
	So, I think it is safe to say that today, the algorithm will be dominated by its memory access pattern, that is why we are also calculating read, and write patterns.
\end{document}