### 2018-05-21 books/FundamentalsMatrixComputations/Exercise.01.005.012.tex

\documentclass{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{pgfplots}
\begin{document}
	Exercise 1.5.12\\
	Consider a mass-spring system as in Figure 1.11 with six carts. Suppose each spring has a stiffness $k_i = 1$ newton/meter.\\ 
	(a) Set up the tridiagonal, positive definite, coefficient matrix $A$ associated with this problem.\\ 
	(b) Use the MATLAB chol command to calculate the Cholesky factor $R$. Notice that R inherits the band structure of A. (To learn an easy way to enter this particular A, type help toeplitz )\\ 
	(c) Use the MATLAB inv command to calculate $A^{-1}$. Notice that $A^{-1}$ does not inherit the band structure of $A$.
	\\
	Answers\\
	\\
	\begin{verbatim}
	octave:1> A = toeplitz([2,-1,0,0,0,0,0])
	A =

	2  -1   0   0   0   0   0
	-1   2  -1   0   0   0   0
	0  -1   2  -1   0   0   0
	0   0  -1   2  -1   0   0
	0   0   0  -1   2  -1   0
	0   0   0   0  -1   2  -1
	0   0   0   0   0  -1   2
	octave:2> chol(A)
	ans =

	1.41421  -0.70711   0.00000   0.00000   0.00000   0.00000   0.00000
	0.00000   1.22474  -0.81650   0.00000   0.00000   0.00000   0.00000
	0.00000   0.00000   1.15470  -0.86603   0.00000   0.00000   0.00000
	0.00000   0.00000   0.00000   1.11803  -0.89443   0.00000   0.00000
	0.00000   0.00000   0.00000   0.00000   1.09545  -0.91287   0.00000
	0.00000   0.00000   0.00000   0.00000   0.00000   1.08012  -0.92582
	0.00000   0.00000   0.00000   0.00000   0.00000   0.00000   1.06904
	octave:3> inv(A)
	ans =

	0.87500   0.75000   0.62500   0.50000   0.37500   0.25000   0.12500
	0.75000   1.50000   1.25000   1.00000   0.75000   0.50000   0.25000
	0.62500   1.25000   1.87500   1.50000   1.12500   0.75000   0.37500
	0.50000   1.00000   1.50000   2.00000   1.50000   1.00000   0.50000
	0.37500   0.75000   1.12500   1.50000   1.87500   1.25000   0.62500
	0.25000   0.50000   0.75000   1.00000   1.25000   1.50000   0.75000
	0.12500   0.25000   0.37500   0.50000   0.62500   0.75000   0.87500
	\end{verbatim}	
\end{document}