### 2018-05-25 books/FundamentalsMatrixComputations/Exercise.01.007.012.tex

\documentclass[10pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\author{Daniel Frederico Lins Leite}
\begin{document}
	We will transform the system by means of elementary operations of three types:\\ 
	1. Add a multiple of one equation to another equation.\\ 
	2. Interchange two equations.\\ 
	3. Multiply an equation by a nonzero constant.\\ 
	\\
	Exercise 1.7.2:\\
	\\
	Proposition 1.7.1\\
	If Ax = b is obtained from Ax = b by an elementary operation of type 1, 2, or 3, then the systems Ax = b and Ax = b are equivalent.\\
	\\
	Exercise 1.7.2\\
	Prove Proposition 1.7.1.\\
	Discussion: Suppose the system $Ax = b$ is transformed to $\hat{A}x = \hat{b}$ by an operation of type 1. You must show that\\
	(a) every solution of $Ax = b$ is a solution of $\hat{A}x = \hat{b}$ and\\
	(b) every solution of $\hat{A}x = \hat{b}$ is a solution of $Ax = b$.\\
	\\
	Part (a) should be easy. Part (b) becomes easy when you realize that $Ax = b$ can be recovered from $\hat{A}x = \hat{b}$ by an operation of type 1: If $\hat{A}x = b$ was obtained from $Ax = b$ by adding m times the j-th row to the i-th row, then $Ax = b$ can be recovered from $\hat{A}x = b$ by adding $-m$ times the j-th row to the i-th row. Analogous remarks apply to operations of types 2 and 3.\\
	\\
	Answers:\\
	\\
	(a)\\
	\\
	\begin{align*}
		Ax = b\\
		\forall{j} \big[ \text{row}_j(A) * x = b_j \big]\\
		\\
		\hat{A} = A + m*\text{row}_j(A)\\
		\hat{b} = b + m*b_j\\	
	\end{align*}
	\begin{align*}	
		Ax = b\\
		Ax + m*\text{row}_j(A)*x = b + m*\text{row}_j(A)*x\\		
		Ax + m*\text{row}_j(A)*x = b + m*b_j\\
		(A + m*\text{row}_j(A))*x = b + m*b_j\\
		(A + m*\text{row}_j(A))*x = \hat{b}\\
		\hat{A}*x = \hat{b}\\		
		\square
	\end{align*} 	
	\\
	(b)\\
	\\
	\begin{align*}
	Ax = b\\
	\forall{j} \big[ \text{row}_j(A) * x = b_j \big]\\
	\\
	\hat{A} = A + m*\text{row}_j(A)\\
	\hat{b} = b + m*b_j\\	
	\\
	\hat{A}x = \hat{b}\\
	(A + m*\text{row}_j(A))*x = b + m*b_j\\
	A*x + m*\text{row}_j(A)*x = b + m*b_j\\
	b + m*\text{row}_j(A)*x = b + m*b_j\\
	m*\text{row}_j(A)*x = b + m*b_j - b\\
	m*\text{row}_j(A)*x = m*b_j\\		
	\text{row}_j(A)*x = b_j\\
	\square	
	\end{align*} 	
\end{document}