### 2018-06-21 books/IntroductionLinearALgebra/Exercise.01.004.tex

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{pdfpages}
\title{2D Graphics with Asymptote}
\author{The Asymptote Project}
\newcommand{\insertrep}[1]{%
	\hspace*{-2.4cm}
	\fbox{\includegraphics[page=1,scale=0.8]{#1}}
}
\begin{document}
Exercise 4:\\
From $v=\begin{bmatrix}2\\1\end{bmatrix}$ and $w=\begin{bmatrix}1\\2\end{bmatrix}$, find the components of $3v+w$ and $cv+dw$.\\
\begin{align*}
	v&=\begin{bmatrix}2\\1\end{bmatrix}\\
	w&=\begin{bmatrix}1\\2\end{bmatrix}\\
	\\
	3v+w &= 3*\begin{bmatrix}2\\1\end{bmatrix} + \begin{bmatrix}1\\2\end{bmatrix}\\
	&=\begin{bmatrix}6\\3\end{bmatrix}+\begin{bmatrix}1\\2\end{bmatrix}\\
	&=\begin{bmatrix}7\\5\end{bmatrix}\\
	\\
	cv+dw &= c*\begin{bmatrix}2\\1\end{bmatrix} + d*\begin{bmatrix}1\\2\end{bmatrix}\\
	&= \begin{bmatrix}2c\\c\end{bmatrix} + \begin{bmatrix}d\\2d\end{bmatrix}\\
	&= \begin{bmatrix}2c+d\\c+2d\end{bmatrix}
\end{align*}
\newpage
Red dots are random linear combinations of $v$ and $w$.
\begin{center}
	\insertrep{exercise01004.pdf}
\end{center}
\end{document}