### 2018-06-21 books/IntroductionLinearALgebra/Exercise.01.006.tex

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{pdfpages}
\title{2D Graphics with Asymptote}
\author{The Asymptote Project}
\newcommand{\insertrep}[1]{%
	\hspace*{-2.4cm}
	\fbox{\includegraphics[page=1,scale=2]{#1}}
}
\begin{document}
Exercise 6:\\
Every combination of $v=(1,-2,1)$ and $w=(0,1,-1)$ has components that  add to? Find $d$ and $d$ so that $cv+dw=(3,3,-6)$.\\
(a)\\
\begin{align*}
	v+w &= \begin{bmatrix}1\\-2\\1\end{bmatrix} + \begin{bmatrix}0\\1\\-1\end{bmatrix}\\
	&= \begin{bmatrix}1\\-1\\0\end{bmatrix}
\end{align*}
(b)
\begin{align*}
cv+dw &= c*\begin{bmatrix}1\\-2\\1\end{bmatrix} + d*\begin{bmatrix}0\\1\\-1\end{bmatrix} = \begin{bmatrix}3\\3\\-6\end{bmatrix}\\
	&= \begin{bmatrix}c*1+d*0\\c*-2+d*1\\c*1+d*-1\end{bmatrix} = \begin{bmatrix}3\\3\\-6\end{bmatrix}\\
	\\
	c*1+d*0&=3\\
	c&=3\\
	\\
	c*-2+d*1 = 3\\
	3*-2+d*1 = 3\\	
	-6+d = 3\\		
	d = 3 + 6\\			
	d = 9\\		
	\\
	c*1+d*-1=-6\\
	3*1+9*-1=-6\\
	3-9=-6\\	
	-6=-6\\
\end{align*}
\end{document}