### 2018-06-24 books/IntroductionLinearALgebra/Exercise.01.011.tex
Introduction to Linear Algebra by GIlbert Strang

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{pdfpages}
\usepackage{lscape}
\title{2D Graphics with Asymptote}
\author{The Asymptote Project}
\newcommand{\insertrep}[1]{%
	\hspace*{-2.4cm}
	\fbox{\includegraphics[page=1,scale=1.5]{#1}}
}
\begin{document}
	Exercise 11:\\
	Four corners of the cube are $(0,0,0)$,$(1,0,0)$,$(0,1,0)$,$(0,0,1)$. What are the other four corners? Find the coordinates of the center point of the cube. The center points of the six faces are?\\
	The other coordinates are the sum of the four known points.\\
	And the center point for each of the cube face's are the averaged sum of itts four corners.
	\begin{center}
		\insertrep{exercise01011.pdf}
	\end{center}
\end{document}