### 2018-06-24 books/IntroductionLinearALgebra/Exercise.01.013.tex
Introduction to Linear Algebra by Gilbert Strang

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{pdfpages}
\usepackage{lscape}
\usepackage{hyperref}
\title{2D Graphics with Asymptote}
\author{The Asymptote Project}
\newcommand{\insertrep}[1]{%
	\hspace*{-2.4cm}
	\fbox{\includegraphics[page=1,scale=0.5]{#1}}
}
\begin{document}
	Exercise 13:\\
	(a) What us the sum V if the twelve vectors that go from the center of a clock to the hours 01:00, 02:00, ..., 12:00?\\
	(b) If the 02:00 vector is removed, why do the 11 remaining vectors add to0 8:00?\\
	(c) What are the components of that 02:00 vector $v = (cos \theta, sin \theta)$?
	\\
	(a) The sum is the zero vector because of the symmetry.\\
	(b) Because 08:00 is the vector that is symmetric to the 02:00 vector.\\
	(c) Each hour is $-\frac{\pi}{12}$, but our zero is actually at 03:00. So 2:00 is $(cos(\frac{2\pi}{12}),sin(\frac{2\pi}{12}))$\\
	\begin{align*}
		cos(\frac{2\pi}{12}) &= 0.866025403784439\\
		sin(\frac{2\pi}{12}) &= 0.5\\
	\end{align*}

	\begin{center}
		\insertrep{exercise01013.pdf}
	\end{center}
\end{document}