### 2018-08-01 books/IntroductionLinearALgebra/Exercise.01.014.tex
.

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{pdfpages}
\usepackage{lscape}
\usepackage{hyperref}
\title{2D Graphics with Asymptote}
\author{The Asymptote Project}
\newcommand{\insertrep}[1]{%
	\hspace*{-2.4cm}
	\fbox{\includegraphics[page=1,scale=0.5]{#1}}
}
\begin{document}
	Exercise 14:\\
	Suppose the twelves vectors start from 06:00 at the bottom instead of $(0,0)$ at the center. The vector to 12:00 is doubled to $(0,2)$. Add the new twelve vectors.\\
	\\
	This is the same as adding $(0,1)$, wich will make the sum result in $(0,12)$ instaead of $(0,0)$.
\end{document}