### 2018-08-01 courses/harvard-applied-calculus/PE.01.tex

\documentclass[10pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\author{Daniel Frederico Lins Leite}
\begin{document}
	Here is the formula for price elasticity between two points $(p_1,q_1)$  and $(p_2,q_2)$:
	\begin{align*}
		\text{Price Elasticity} = \frac{\triangle q}{q_1} / \frac{\triangle p}{p_1}
	\end{align*}		
	Rearrange the formula and use the assumption that  to write a formula for price elasticity of demand at the point . Select the correct formula below.				
	\\
	Answer:\\
	\begin{align*}
		 \frac{\triangle q}{q_1} / \frac{\triangle p}{p_1}  &= \frac{\triangle q}{q_1} * \frac{p_1}{\triangle p}\\
		 &= \frac{\triangle q}{\triangle p} * \frac{p_1}{q_1}\\
 		 &= \frac{dq}{dp}\big|_{p_1} * \frac{p_1}{q_1}
	\end{align*}			
\end{document}